---
title: "R Notebook"
output: html_notebook
---

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

```{r}
# Import necessary packages and set working directory
library(tidyverse)
library(readxl)
library(lubridate)
library(baseballr)

setwd('~/R/Projects/Syndergaard_Project')

# Create vector of pitcher names and look up IDs
pitchers <- c('Syndergaard', 'Verlander', 'Bauer', 'deGrom', 'Bieber')
for (p in pitchers) {
  print(playerid_lookup(p))
}
pitcher_ids <- c(592789, 434378, 545333, 594798, 669456)
```

```{r}
# Syndergaard
# Scrape Savant to get Pitch Movement for Break Chart
syndergaard_savant_scrape2015 <- scrape_statcast_savant(start_date = "2015-03-01",
                                                        playerid = pitcher_ids[1], player_type='pitcher')
syndergaard_savant_scrape2016 <- scrape_statcast_savant(start_date = "2016-03-01",
                                                        playerid = pitcher_ids[1], player_type='pitcher')
syndergaard_savant_scrape2017 <- scrape_statcast_savant(start_date = "2017-03-01",
                                                        playerid = pitcher_ids[1], player_type='pitcher')
syndergaard_savant_scrape2018 <- scrape_statcast_savant(start_date = "2018-03-01",
                                                        playerid = pitcher_ids[1], player_type='pitcher')
syndergaard_savant_scrape2019 <- scrape_statcast_savant(start_date = "2019-03-01",
                                                        playerid = pitcher_ids[1], player_type='pitcher')

# Combine
syndergaard_savant_scrape <- rbind(syndergaard_savant_scrape2015,
                                   syndergaard_savant_scrape2016, 
                                   syndergaard_savant_scrape2017,
                                   syndergaard_savant_scrape2018, 
                                   syndergaard_savant_scrape2019)
syndergaard_savant_scrape$game_year <- year(as.Date(as.character(syndergaard_savant_scrape$game_year), format = '%Y'))
syndergaard_savant_scrape$pitch_type <- as.factor(syndergaard_savant_scrape$pitch_type)
syndergaard_savant_scrape$pitch_name <- as.factor(syndergaard_savant_scrape$pitch_name)
syndergaard_savant_scrape$pitch_type <- syndergaard_savant_scrape$pitch_type %>% plyr::revalue(c('FT' = 'SI'))
syndergaard_savant_scrape$pitch_name <- syndergaard_savant_scrape$pitch_name %>% plyr::revalue(c('2-Seam Fastball' = 'Sinker'))
syndergaard_savant_scrape[, c(70, 71, 72, 73, 74, 75, 76)] <- sapply(syndergaard_savant_scrape[, c(70, 71, 72, 73, 74, 75, 76)], as.numeric)

# Put movement in inches
syndergaard_savant_scrape <- syndergaard_savant_scrape %>% 
  mutate(pfx_x = round(pfx_x * 12, 1), pfx_z = round(pfx_z * 12, 1))

# Remove extraneous pitch types
syndergaard_savant_scrape <- syndergaard_savant_scrape %>%
  filter(pitch_type != 'PO' & pitch_type != 'IN' & pitch_type != 'null') %>%
  droplevels()

# Identify extraneous columns
colnames(syndergaard_savant_scrape)

# Select only necessary columns
syndergaard_savant_scrape <- syndergaard_savant_scrape %>% 
         select(1:10, 15, 24:26, 27:31, 54:58, 71:73, 76:79)

# Create Syndergaard pfx DataFrame
syndergaard_pfx <- syndergaard_savant_scrape %>% 
  group_by(game_year, player_name, pitch_type, pitch_name) %>%
  summarize(pitches = n(),
            pfx_x = round(mean(pfx_x), 1), 
            pfx_z = round(mean(pfx_z), 1), 
            speed = round(mean(release_speed, na.rm = TRUE), 1),
            spin_rate = round(mean(release_spin_rate, na.rm = TRUE)),
            bauer_units = round(spin_rate/speed, 1))

syndergaard_pfx <- syndergaard_pfx %>% group_by(game_year) %>% mutate(total_pitches = sum(pitches))
syndergaard_pfx <- syndergaard_pfx %>% mutate(percent_usage = round(pitches / total_pitches * 100, 1))
syndergaard_pfx <- syndergaard_pfx[, c(1:5, 11:12, 6:10)]
syndergaard_pfx
```

```{r}
# Verlander
# Scrape Savant to get Pitch Movement for Break Chart
verlander_savant_scrape2015 <- scrape_statcast_savant(start_date = "2015-03-01",
                                                        playerid = pitcher_ids[2], player_type='pitcher')
verlander_savant_scrape2016 <- scrape_statcast_savant(start_date = "2016-03-01",
                                                        playerid = pitcher_ids[2], player_type='pitcher')
verlander_savant_scrape2017 <- scrape_statcast_savant(start_date = "2017-03-01",
                                                        playerid = pitcher_ids[2], player_type='pitcher')
verlander_savant_scrape2018 <- scrape_statcast_savant(start_date = "2018-03-01",
                                                        playerid = pitcher_ids[2], player_type='pitcher')
verlander_savant_scrape2019 <- scrape_statcast_savant(start_date = "2019-03-01",
                                                        playerid = pitcher_ids[2], player_type='pitcher')
verlander_savant_scrape2020 <- scrape_statcast_savant(start_date = "2020-03-01",
                                                        playerid = pitcher_ids[2], player_type='pitcher')

# Combine
verlander_savant_scrape <- rbind(verlander_savant_scrape2015,
                                 verlander_savant_scrape2016, 
                                 verlander_savant_scrape2017,
                                 verlander_savant_scrape2018, 
                                 verlander_savant_scrape2019,
                                 verlander_savant_scrape2020)
verlander_savant_scrape$game_year <- year(as.Date(as.character(verlander_savant_scrape$game_year), format = '%Y'))
verlander_savant_scrape$pitch_type <- as.factor(verlander_savant_scrape$pitch_type)
verlander_savant_scrape[, c(70, 71, 72, 73, 74, 75, 76)] <- sapply(verlander_savant_scrape[, c(70, 71, 72, 73, 74, 75, 76)], as.numeric)

# Put movement in inches
verlander_savant_scrape <- verlander_savant_scrape %>% 
  mutate(pfx_x = round(pfx_x * 12, 1), pfx_z = round(pfx_z * 12, 1))

# Remove extraneous pitch types
verlander_savant_scrape <- verlander_savant_scrape %>%
  filter(pitch_type != 'PO' & pitch_type != 'IN' & pitch_type != 'null') %>%
  droplevels()

# Identify extraneous columns
colnames(verlander_savant_scrape)

# Select only necessary columns
verlander_savant_scrape <- verlander_savant_scrape %>% 
         select(1:10, 15, 24:26, 27:31, 54:58, 71:73, 76:79)

# Create Verlander pfx DataFrame
verlander_pfx <- verlander_savant_scrape %>% 
  group_by(game_year, player_name, pitch_type, pitch_name) %>%
  summarize(pitches = n(),
            pfx_x = round(mean(pfx_x), 1), 
            pfx_z = round(mean(pfx_z), 1), 
            speed = round(mean(release_speed, na.rm = TRUE), 1),
            spin_rate = round(mean(release_spin_rate, na.rm = TRUE)),
            bauer_units = round(spin_rate/speed, 1))
verlander_pfx <- verlander_pfx %>% group_by(game_year) %>% mutate(total_pitches = sum(pitches))
verlander_pfx <- verlander_pfx %>% mutate(percent_usage = round(pitches / total_pitches * 100, 1))
verlander_pfx <- verlander_pfx[, c(1:5, 11:12, 6:10)]
verlander_pfx
```
```{r}
# Bauer
# Scrape Savant to get Pitch Movement for Break Chart
bauer_savant_scrape2015 <- scrape_statcast_savant(start_date = "2015-03-01",
                                                        playerid = pitcher_ids[3], player_type='pitcher')
bauer_savant_scrape2016 <- scrape_statcast_savant(start_date = "2016-03-01",
                                                        playerid = pitcher_ids[3], player_type='pitcher')
bauer_savant_scrape2017 <- scrape_statcast_savant(start_date = "2017-03-01",
                                                        playerid = pitcher_ids[3], player_type='pitcher')
bauer_savant_scrape2018 <- scrape_statcast_savant(start_date = "2018-03-01",
                                                        playerid = pitcher_ids[3], player_type='pitcher')
bauer_savant_scrape2019 <- scrape_statcast_savant(start_date = "2019-03-01",
                                                        playerid = pitcher_ids[3], player_type='pitcher')
bauer_savant_scrape2020 <- scrape_statcast_savant(start_date = "2020-03-01",
                                                        playerid = pitcher_ids[3], player_type='pitcher')

# Combine
bauer_savant_scrape <- rbind(bauer_savant_scrape2015,
                             bauer_savant_scrape2016, 
                             bauer_savant_scrape2017,
                             bauer_savant_scrape2018, 
                             bauer_savant_scrape2019,
                             bauer_savant_scrape2020)
bauer_savant_scrape$game_year <- year(as.Date(as.character(bauer_savant_scrape$game_year), format = '%Y'))
bauer_savant_scrape$pitch_name <- as.factor(bauer_savant_scrape$pitch_name)
bauer_savant_scrape$pitch_name <- bauer_savant_scrape$pitch_name %>% plyr::revalue(c('Knuckle Curve' = 'Curveball'))
bauer_savant_scrape$pitch_type <- as.factor(bauer_savant_scrape$pitch_type)
bauer_savant_scrape$pitch_type <- bauer_savant_scrape$pitch_type %>% plyr::revalue(c('KC' = 'CU'))
bauer_savant_scrape[, c(70, 71, 72, 73, 74, 75, 76)] <- sapply(bauer_savant_scrape[, c(70, 71, 72, 73, 74, 75, 76)], as.numeric)

# Put movement in inches
bauer_savant_scrape <- bauer_savant_scrape %>% 
  mutate(pfx_x = round(pfx_x * 12, 1), pfx_z = round(pfx_z * 12, 1))

# Remove extraneous pitch types
bauer_savant_scrape <- bauer_savant_scrape %>%
  filter(pitch_type != 'PO' & pitch_type != 'IN' & pitch_type != 'null') %>%
  droplevels()

# Identify extraneous columns
colnames(bauer_savant_scrape)

# Select only necessary columns
bauer_savant_scrape <- bauer_savant_scrape %>% 
         select(1:10, 15, 24:26, 27:31, 54:58, 71:73, 76:79)

# Create Bauer pfx DataFrame
bauer_pfx <- bauer_savant_scrape %>% 
  group_by(game_year, player_name, pitch_type, pitch_name) %>%
  summarize(pitches = n(),
            pfx_x = round(mean(pfx_x), 1), 
            pfx_z = round(mean(pfx_z), 1), 
            speed = round(mean(release_speed, na.rm = TRUE), 1),
            spin_rate = round(mean(release_spin_rate, na.rm = TRUE)),
            bauer_units = round(spin_rate/speed, 1))
bauer_pfx <- bauer_pfx %>% group_by(game_year) %>% mutate(total_pitches = sum(pitches))
bauer_pfx <- bauer_pfx %>% mutate(percent_usage = round(pitches / total_pitches * 100, 1))
bauer_pfx <- bauer_pfx[, c(1:5, 11:12, 6:10)]
bauer_pfx
```
```{r}
# deGrom
# Scrape Savant to get Pitch Movement for Break Chart
degrom_savant_scrape2015 <- scrape_statcast_savant(start_date = "2015-03-01",
                                                        playerid = pitcher_ids[4], player_type='pitcher')
degrom_savant_scrape2016 <- scrape_statcast_savant(start_date = "2016-03-01",
                                                        playerid = pitcher_ids[4], player_type='pitcher')
degrom_savant_scrape2017 <- scrape_statcast_savant(start_date = "2017-03-01",
                                                        playerid = pitcher_ids[4], player_type='pitcher')
degrom_savant_scrape2018 <- scrape_statcast_savant(start_date = "2018-03-01",
                                                        playerid = pitcher_ids[4], player_type='pitcher')
degrom_savant_scrape2019 <- scrape_statcast_savant(start_date = "2019-03-01",
                                                        playerid = pitcher_ids[4], player_type='pitcher')
degrom_savant_scrape2020 <- scrape_statcast_savant(start_date = "2020-03-01",
                                                        playerid = pitcher_ids[4], player_type='pitcher')

# Combine
degrom_savant_scrape <- rbind(degrom_savant_scrape2015,
                              degrom_savant_scrape2016, 
                              degrom_savant_scrape2017,
                              degrom_savant_scrape2018, 
                              degrom_savant_scrape2019,
                              degrom_savant_scrape2020)
degrom_savant_scrape$game_year <- year(as.Date(as.character(degrom_savant_scrape$game_year), format = '%Y'))
degrom_savant_scrape$pitch_type <- as.factor(degrom_savant_scrape$pitch_type)
degrom_savant_scrape[, c(70, 71, 72, 73, 74, 75, 76)] <- sapply(degrom_savant_scrape[, c(70, 71, 72, 73, 74, 75, 76)], as.numeric)

# Put movement in inches
degrom_savant_scrape <- degrom_savant_scrape %>% 
  mutate(pfx_x = round(pfx_x * 12, 1), pfx_z = round(pfx_z * 12, 1))

# Remove extraneous pitch types
degrom_savant_scrape <- degrom_savant_scrape %>%
  filter(pitch_type != 'PO' & pitch_type != 'IN' & pitch_type != 'null') %>%
  droplevels()

# Identify extraneous columns
colnames(degrom_savant_scrape)

# Select only necessary columns
degrom_savant_scrape <- degrom_savant_scrape %>% 
         select(1:10, 15, 24:26, 27:31, 54:58, 71:73, 76:79)

# Create deGrom pfx DataFrame
degrom_pfx <- degrom_savant_scrape %>% 
  group_by(game_year, player_name, pitch_type, pitch_name) %>%
  summarize(pitches = n(),
            pfx_x = round(mean(pfx_x), 1), 
            pfx_z = round(mean(pfx_z), 1), 
            speed = round(mean(release_speed, na.rm = TRUE), 1),
            spin_rate = round(mean(release_spin_rate, na.rm = TRUE)),
            bauer_units = round(spin_rate/speed, 1))
degrom_pfx <- degrom_pfx %>% group_by(game_year) %>% mutate(total_pitches = sum(pitches))
degrom_pfx <- degrom_pfx %>% mutate(percent_usage = round(pitches / total_pitches * 100, 1))
degrom_pfx <- degrom_pfx[, c(1:5, 11:12, 6:10)]
degrom_pfx
```
```{r}
# Bieber
# Scrape Savant to get Pitch Movement for Break Chart
bieber_savant_scrape2018 <- scrape_statcast_savant(start_date = "2018-03-01",
                                                        playerid = pitcher_ids[5], player_type='pitcher')
bieber_savant_scrape2019 <- scrape_statcast_savant(start_date = "2019-03-01",
                                                        playerid = pitcher_ids[5], player_type='pitcher')
bieber_savant_scrape2020 <- scrape_statcast_savant(start_date = "2020-03-01",
                                                        playerid = pitcher_ids[5], player_type='pitcher')

# Combine
bieber_savant_scrape <- rbind(bieber_savant_scrape2018, 
                              bieber_savant_scrape2019,
                              bieber_savant_scrape2020)
bieber_savant_scrape$game_year <- year(as.Date(as.character(bieber_savant_scrape$game_year), format = '%Y'))
bieber_savant_scrape$pitch_name <- as.factor(bieber_savant_scrape$pitch_name)
bieber_savant_scrape$pitch_name <- bieber_savant_scrape$pitch_name %>% plyr::revalue(c('Knuckle Curve' = 'Curveball'))
bieber_savant_scrape$pitch_type <- as.factor(bieber_savant_scrape$pitch_type)
bieber_savant_scrape$pitch_type <- bieber_savant_scrape$pitch_type %>% plyr::revalue(c('KC' = 'CU'))
bieber_savant_scrape[, c(70, 71, 72, 73, 74, 75, 76)] <- sapply(bieber_savant_scrape[, c(70, 71, 72, 73, 74, 75, 76)], as.numeric)

# Put movement in inches
bieber_savant_scrape <- bieber_savant_scrape %>% 
  mutate(pfx_x = round(pfx_x * 12, 1), pfx_z = round(pfx_z * 12, 1))

# Remove extraneous pitch types
bieber_savant_scrape <- bieber_savant_scrape %>%
  filter(pitch_type != 'PO' & pitch_type != 'IN' & pitch_type != 'null') %>%
  droplevels()

# Identify extraneous columns
colnames(bieber_savant_scrape)

# Select only necessary columns
bieber_savant_scrape <- bieber_savant_scrape %>% 
         select(1:10, 15, 24:26, 27:31, 54:58, 71:73, 76:79)

# Create Bieber pfx DataFrame
bieber_pfx <- bieber_savant_scrape %>% 
  group_by(game_year, player_name, pitch_type, pitch_name) %>%
  summarize(pitches = n(),
            pfx_x = round(mean(pfx_x), 1), 
            pfx_z = round(mean(pfx_z), 1), 
            speed = round(mean(release_speed, na.rm = TRUE), 1),
            spin_rate = round(mean(release_spin_rate, na.rm = TRUE)),
            bauer_units = round(spin_rate/speed, 1))
bieber_pfx <- bieber_pfx %>% group_by(game_year) %>% mutate(total_pitches = sum(pitches))
bieber_pfx <- bieber_pfx %>% mutate(percent_usage = round(pitches / total_pitches * 100, 1))
bieber_pfx <- bieber_pfx[, c(1:5, 11:12, 6:10)]
bieber_pfx
```

```{r}
# Combine pfx DataFrames
pitchers_fx <- rbind(syndergaard_pfx,
                     verlander_pfx,
                     bauer_pfx,
                     degrom_pfx,
                     bieber_pfx)

# Read in Savant Data (wOBA, xwOBA, etc.)
savant_data <- read_csv('savant_data_2015-2020.csv')
savant_data <- savant_data %>% select(1:23)

# Add whiff rate and change column type as necessary
savant_data <- savant_data %>% mutate(whiff_rate = round(whiffs/swings,3) * 100)
savant_data$pitch_type <- as.factor(savant_data$pitch_type)
savant_data[, c(8:18)] <- sapply(savant_data[, c(8:18)], as.numeric)

# Reorder columns
savant_data <- savant_data[, c(4, 1, 5, 2, 3, 6, 7, 8:24)]
savant_data

# Attach pitch names to Savant chart
savant_data <- merge(savant_data, pitchers_fx[, c('game_year', 'player_name', 'pitch_type', 'pitch_name')], 
                     by = c('game_year', 'player_name', 'pitch_type'), all.x = TRUE)
```

```{r}
# Charts
# Create Grid to make Breaks clearer
topGrid <- 30
botGrid <- -30
centerGrid <- 0
rightGrid <- 30
leftGrid <- -30
Grid <- data.frame(
  x=c(centerGrid, centerGrid, centerGrid, leftGrid, rightGrid),
  y=c(topGrid, botGrid, centerGrid, centerGrid, centerGrid)
)

# Syndergaard Break Chart
syndergaard_break_chart <- syndergaard_savant_scrape %>% 
  ggplot(aes(pfx_x, pfx_z, color = pitch_name)) + geom_point() + 
  geom_path(aes(x, y), data=Grid, lwd=2, col="black") + 
  ylim(-30, 30) + xlim(-30, 30) + xlab('Horizontal Movement (in.)') + ylab('Vertical Movement (in.)') + 
  ggtitle('Syndergaard Break Charts by Year') +
  facet_wrap(~ game_year)

syndergaard_break_chart
```
```{r}
# Syndergaard Average Break Chart
syndergaard_break_chart_avg <- syndergaard_pfx %>% 
  ggplot(aes(pfx_x, pfx_z, color = pitch_name)) + geom_point(shape = 20, size = 7) + 
  geom_path(aes(x, y), data=Grid, lwd=2, col="black") + 
  ylim(-30, 30) + xlim(-30, 30) + xlab('Horizontal Movement (in.)') + ylab('Vertical Movement (in.)') + 
  ggtitle('Syndergaard  Break Charts by Year') +
  facet_wrap(~ game_year)

syndergaard_break_chart_avg
```

```{r}
# Verlander Break Chart 2016
verlander_break_chart_2016 <- verlander_pfx %>%
  filter(game_year == 2016) %>%
  ggplot(aes(pfx_x, pfx_z, color = pitch_name)) + geom_point(shape = 20, size = 7) + 
  geom_path(aes(x, y), data=Grid, lwd=2, col="black") + 
  ylim(-30, 30) + xlim(-30, 30) + xlab('Horizontal Movement (in.)') + ylab('Vertical Movement (in.)') + 
  ggtitle('Verlander Average Break Chart 2016') 

verlander_break_chart_2016
```

```{r}
# Verlander Break Chart early 2017
verlander_break_chart_early_2017 <- verlander_savant_scrape %>%
  filter(game_date > '2017-03-01' & game_date < '2017-06-21') %>%
  ggplot(aes(pfx_x, pfx_z, color = pitch_name)) + geom_point() + 
  geom_path(aes(x, y), data=Grid, lwd=2, col="black") + 
  ylim(-30, 30) + xlim(-30, 30) + xlab('Horizontal Movement (in.)') + ylab('Vertical Movement (in.)') + 
  ggtitle('Verlander Break Chart early 2017 (prior to June 21st)') 

verlander_break_chart_early_2017
```
```{r}
# Verlander Average Break Chart early 2017
verlander_break_chart_early_2017_avg <- verlander_savant_scrape %>%
  filter(game_date > '2017-03-01' & game_date < '2017-06-21') %>%
  group_by(pitch_name) %>%
  summarize(pfx_x = mean(pfx_x), pfx_z = mean(pfx_z)) %>%
  ggplot(aes(pfx_x, pfx_z, color = pitch_name)) + geom_point(shape = 20, size = 7) + 
  geom_path(aes(x, y), data=Grid, lwd=2, col="black") + 
  ylim(-30, 30) + xlim(-30, 30) + xlab('Horizontal Movement (in.)') + ylab('Vertical Movement (in.)') + 
  ggtitle('Verlander Break Chart early 2017 (prior to June 21st)') 

verlander_break_chart_early_2017_avg
```

```{r}
# Verlander Break Chart late 2017
verlander_break_chart_late_2017 <- verlander_savant_scrape %>%
  filter(game_date >= '2017-06-21' & game_date < '2017-11-01') %>%
  ggplot(aes(pfx_x, pfx_z, color = pitch_name)) + geom_point() + 
  geom_path(aes(x, y), data=Grid, lwd=2, col="black") + 
  ylim(-30, 30) + xlim(-30, 30) + xlab('Horizontal Movement (in.)') + ylab('Vertical Movement (in.)') + 
  ggtitle('Verlander Break Chart late 2017 (after June 21st)') 

verlander_break_chart_late_2017
```
```{r}
# Verlander Average Break Chart late 2017
verlander_break_chart_late_2017_avg <- verlander_savant_scrape %>%
  filter(game_date >= '2017-06-21' & game_date < '2017-11-01') %>%
  group_by(pitch_name) %>%
  summarize(pfx_x = mean(pfx_x), pfx_z = mean(pfx_z)) %>%
  ggplot(aes(pfx_x, pfx_z, color = pitch_name)) + geom_point(shape = 20, size = 7) + 
  geom_path(aes(x, y), data=Grid, lwd=2, col="black") + 
  ylim(-30, 30) + xlim(-30, 30) + xlab('Horizontal Movement (in.)') + ylab('Vertical Movement (in.)') + 
  ggtitle('Verlander Break Chart late 2017 (after to June 21st)') 

verlander_break_chart_late_2017_avg
```

```{r}
# Verlander Break Chart 2019
verlander_break_chart <- syndergaard_savant_scrape %>% 
  filter(game_year == 2019) %>%
  ggplot(aes(pfx_x, pfx_z, color = pitch_name)) + geom_point() + 
  geom_path(aes(x, y), data=Grid, lwd=2, col="black") + 
  ylim(-30, 30) + xlim(-30, 30) + xlab('Horizontal Movement (in.)') + ylab('Vertical Movement (in.)') + 
  ggtitle('Verlander Break Chart 2019')

verlander_break_chart
```

```{r}
# Verlander Average Break Chart 2019
verlander_break_chart_2019_avg <- verlander_pfx %>%
  filter(game_year == 2019) %>%
  ggplot(aes(pfx_x, pfx_z, color = pitch_name)) + geom_point(shape = 20, size = 7) + 
  geom_path(aes(x, y), data=Grid, lwd=2, col="black") + 
  ylim(-30, 30) + xlim(-30, 30) + xlab('Horizontal Movement (in.)') + ylab('Vertical Movement (in.)') + 
  ggtitle('Verlander Break Chart 2019') 

verlander_break_chart_2019_avg
```

```{r}
# Bauer Break Chart
bauer_break_chart <- bauer_savant_scrape %>%
  filter(game_year == 2020) %>%
  ggplot(aes(pfx_x, pfx_z, color = pitch_name)) + geom_point() + 
  geom_path(aes(x, y), data=Grid, lwd=2, col="black") + 
  ylim(-30, 30) + xlim(-30, 30) + xlab('Horizontal Movement (in.)') + ylab('Vertical Movement (in.)') + 
  ggtitle('Bauer Break Chart 2020')

bauer_break_chart
```

```{r}
# Bauer Average Break Chart
bauer_break_chart_avg <- bauer_pfx %>%
  filter(game_year == 2020) %>%
  ggplot(aes(pfx_x, pfx_z, color = pitch_name)) + geom_point(shape = 20, size = 7) + 
  geom_path(aes(x, y), data=Grid, lwd=2, col="black") + 
  ylim(-30, 30) + xlim(-30, 30) + xlab('Horizontal Movement (in.)') + ylab('Vertical Movement (in.)') + 
  ggtitle('Bauer Break Chart 2020')

bauer_break_chart_avg
```

```{r}
# Bieber Break Chart
bieber_break_chart <- bieber_savant_scrape %>%
  filter(game_year == 2020) %>%
  ggplot(aes(pfx_x, pfx_z, color = pitch_name)) + geom_point() + 
  geom_path(aes(x, y), data=Grid, lwd=2, col="black") + 
  ylim(-30, 30) + xlim(-30, 30) + xlab('Horizontal Movement (in.)') + ylab('Vertical Movement (in.)') + 
  ggtitle('Bieber Break Chart 2020')

bieber_break_chart
```
```{r}
# Bieber Average Break Chart 2019
bieber_break_chart_avg <- bieber_pfx %>%
  filter(game_year == 2019) %>%
  ggplot(aes(pfx_x, pfx_z, color = pitch_name)) + geom_point(shape = 20, size = 7) + 
  geom_path(aes(x, y), data=Grid, lwd=2, col="black") + 
  ylim(-30, 30) + xlim(-30, 30) + xlab('Horizontal Movement (in.)') + ylab('Vertical Movement (in.)') + 
  ggtitle('Bieber Break Chart 2020')

bieber_break_chart_avg
```

```{r}
# Bieber Average Break Chart 2020
bieber_break_chart_avg <- bieber_pfx %>%
  filter(game_year == 2020) %>%
  ggplot(aes(pfx_x, pfx_z, color = pitch_name)) + geom_point(shape = 20, size = 7) + 
  geom_path(aes(x, y), data=Grid, lwd=2, col="black") + 
  ylim(-30, 30) + xlim(-30, 30) + xlab('Horizontal Movement (in.)') + ylab('Vertical Movement (in.)') + 
  ggtitle('Bieber Break Chart 2020')

bieber_break_chart_avg
```

```{r}
# deGrom Break Charts from 2015-2020
degrom_break_chart <- degrom_savant_scrape %>%
  ggplot(aes(pfx_x, pfx_z, color = pitch_name)) + geom_point() + 
  geom_path(aes(x, y), data=Grid, lwd=2, col="black") + 
  ylim(-30, 30) + xlim(-30, 30) + xlab('Horizontal Movement (in.)') + ylab('Vertical Movement (in.)') + 
  ggtitle('deGrom Break Charts by Year') +
  facet_wrap(~ game_year)

degrom_break_chart
```
```{r}
# deGrom Average Break Charts from 2015-2020
degrom_break_chart_avg <- degrom_pfx %>%
  ggplot(aes(pfx_x, pfx_z, color = pitch_name)) + geom_point(shape = 20, size = 7) + 
  geom_path(aes(x, y), data=Grid, lwd=2, col="black") + 
  ylim(-30, 30) + xlim(-30, 30) + xlab('Horizontal Movement (in.)') + ylab('Vertical Movement (in.)') + 
  ggtitle('deGrom Break Charts by Year') +
  facet_wrap(~ game_year)

degrom_break_chart_avg
```

```{r}
# Graph vertical pitch movement over career
syndergaard_vert_movement <- syndergaard_pfx %>% 
  ggplot(aes(x = game_year, y = pfx_z, color = pitch_name, group = pitch_name)) +
  geom_line() +
  geom_point(shape = 20, size = 5) +
  xlab("Year") +
  ylab('Vertical Movement') +
  ggtitle('Syndergaard Vertical Movement Throughout Career')

syndergaard_vert_movement
```
```{r}
# Graph vertical pitch movement over career minus 2017
amended_syndergaard_vert_movement <- syndergaard_pfx %>%
  filter(game_year != 2017) %>%
  ggplot(aes(x = game_year, y = pfx_z, color = pitch_name, group = pitch_name)) +
  geom_line() +
  geom_point(shape = 20, size = 5) +
  xlab("Year") +
  ylab('Vertical Movement') +
  ggtitle('Syndergaard Vertical Movement Throughout Career (Minus 2017)')

amended_syndergaard_vert_movement
```


```{r}
# Graph horizontal pitch movement over career
syndergaard_horiz_movement <- syndergaard_pfx %>% 
  ggplot(aes(x = game_year, y = pfx_x, color = pitch_name, group = pitch_name)) +
  geom_line() +
  geom_point(shape = 20, size = 5) +
  xlab("Year") +
  ylab('Horizontal Movement') +
  ggtitle('Syndergaard Horizontal Movement Throughout Career')

syndergaard_horiz_movement
```

```{r}
# Graph horizontal pitch movement over career minus 2017
amended_syndergaard_horiz_movement <- syndergaard_pfx %>%
  filter(game_year != 2017) %>%
  ggplot(aes(x = game_year, y = pfx_x, color = pitch_name, group = pitch_name)) +
  geom_line() +
  geom_point(shape = 20, size = 5) +
  xlab("Year") +
  ylab('Horizontal Movement') +
  ggtitle('Syndergaard Horizontal Movement Throughout Career (Minus 2017)')

amended_syndergaard_horiz_movement
```

```{r}
# Syndergaard Velocity by year
# Syndergaard
syndergaard_velocities <- ggplot(syndergaard_pfx, aes(x = game_year, y = speed, color = pitch_name, group = pitch_name)) +
  geom_line() +
  geom_point(shape = 20, size = 5) +
  xlab('Year') +
  ylab('Velocities') +
  ggtitle('Syndergaard Velocities by Pitch & Year')

syndergaard_velocities
```

```{r}
# Syndergaard Spin Rate by year
syndergaard_spin_rates <- syndergaard_pfx %>%
  ggplot(aes(x = game_year, y = spin_rate, color = pitch_name, group = pitch_name)) +
  geom_line() +
  geom_point(shape = 20, size = 5) +
  xlab('Year') +
  ylab('Spin Rates') +
  ggtitle('Syndergaard Spin Rates by Pitch & Year')

syndergaard_spin_rates
```
```{r}
# Syndergaard Spin Rate by year
syndergaard_ff_spin_rate <- syndergaard_pfx %>%
  filter(pitch_type == 'FF') %>%
  ggplot(aes(x = game_year, y = spin_rate)) +
  geom_line(color = 'orange') +
  geom_point(shape = 20, size = 5, color = 'orange') +
  ylim(1600, 2300) +
  xlab('Year') +
  ylab('Spin Rates') +
  ggtitle('Syndergaard Fastball Spin Rates Throughout Career')

syndergaard_ff_spin_rate
```

```{r}
# Bauer Units Comparison
bauer_units_by_year <- pitchers_fx %>%
  filter(pitch_type == 'FF') %>%
  ggplot(aes(x = game_year, y = bauer_units, color = player_name, group = player_name)) +
  geom_line() +
  geom_point(shape = 20, size = 5) +
  xlab('Year') +
  ylab('Bauer Units') +
  ggtitle("Bauer Units for Each Pitcher's 4-Seam Fastball for last Six Seasons") 

bauer_units_by_year
```

```{r}
# Usage Rates
# Syndergaard
syndergaard_pitch_mix <- ggplot(syndergaard_pfx, aes(x = game_year, y = percent_usage, color = pitch_name, group = pitch_name))+
  geom_line() +
  geom_point(shape = 20, size = 5) +
  xlab('Year') +
  ylab('Percent of Pitch Mix') +
  ggtitle('Syndergaard Pitch Mix by Year')

syndergaard_pitch_mix
```

```{r}
# Usage Rates
# Syndergaard
syndergaard_pitch_mix_alt <- syndergaard_pfx %>%
  filter(pitch_type != 'FT') %>%
  ggplot(aes(x = percent_usage, y = pitch_name, fill = pitch_name)) +
  geom_bar(stat = 'identity', show.legend = FALSE) +
  xlab('Percent of Pitch Mix') + 
  ylab('Pitch Type') +
  ggtitle('Syndergaard Pitch Selection by Year') +
  facet_wrap(~ game_year)

syndergaard_pitch_mix_alt
```

```{r}
degrom_pitch_mix_alt <- degrom_pfx %>%
  ggplot(aes(x = percent_usage, y = pitch_name, fill = pitch_name)) +
  geom_bar(stat = 'identity', show.legend = FALSE) +
  xlab('Percent of Pitch Mix') + 
  ylab('Pitch Type') +
  ggtitle('deGrom Pitch Selection by Year') +
  facet_wrap(~ game_year)

degrom_pitch_mix_alt
```

```{r}
# Syndergaard Whiff Rates by year
syndergaard_whiff_rates <- savant_data %>%
  filter(player_name == 'Noah Syndergaard') %>%
  ggplot(aes(x = game_year, y = whiff_rate, color = pitch_name, group = pitch_name)) +
  geom_line() +
  geom_point(shape = 20, size = 5) +
  xlab('Year') +
  ylab('Whiff Rates') +
  ggtitle('Syndergaard Whiff Rates by Pitch and Year')

syndergaard_whiff_rates
```

```{r}
# Syndergaard plot of fastball location over years
# Create K Zone to overlay onto pitch charts
topKzone <- 3.5 
botKzone <- 1.6 
inKzone <- -0.95 
outKzone <- 0.95 
kZone <- data.frame(
  x=c(inKzone, inKzone, outKzone, outKzone, inKzone),
  y=c(botKzone, topKzone, topKzone, botKzone, botKzone)
)

syndergaard_ff_loc <- syndergaard_savant_scrape %>%
  filter(pitch_type == 'FF') %>%
  ggplot(aes(plate_x, plate_z)) + 
  geom_hex(bins = 15) + 
  scale_fill_continuous(type = "viridis") +
  theme_bw() +
  geom_path(aes(x, y), data=kZone, lwd=2, col="red") + 
  xlab('Horizontal Pitch Location') +
  ylab('Vertical Pitch Location') + 
  facet_wrap(~ game_year) +
  ggtitle('Syndergaard Fastball Location by Year')

syndergaard_ff_loc
```

```{r}
syndergaard_ff_loc_alt <- syndergaard_savant_scrape %>%
  filter(pitch_type == 'FF') %>%
  ggplot(aes(plate_x, plate_z)) + 
  geom_density_2d_filled(contour_var = "ndensity") +
  geom_path(aes(x, y), data=kZone, lwd=2, col="red") + 
  xlab('Horizontal Pitch Location') +
  ylab('Vertical Pitch Location') + 
  facet_wrap(~ game_year) +
  ggtitle('Syndergaard Fastball Location by Year')

syndergaard_ff_loc_alt
```

```{r}
degrom_ff_loc <- degrom_savant_scrape %>%
  filter(pitch_type == 'FF') %>%
  ggplot(aes(plate_x, plate_z)) + 
  geom_hex(bins = 15) + 
  scale_fill_continuous(type = "viridis") +
  theme_bw() +
  geom_path(aes(x, y), data=kZone, lwd=2, col="red") + 
  xlab('Horizontal Pitch Location') +
  ylab('Vertical Pitch Location') +
  ylim(0, 5) +
  facet_wrap(~ game_year) +
  ggtitle('deGrom Fastball Location by Year')

degrom_ff_loc
```

```{r}
degrom_ff_loc_alt <- degrom_savant_scrape %>%
  filter(pitch_type == 'FF') %>%
  ggplot(aes(plate_x, plate_z)) + 
  geom_density_2d_filled(contour_var = "ndensity") +
  geom_path(aes(x, y), data=kZone, lwd=2, col="red") + 
  xlab('Horizontal Pitch Location') +
  ylab('Vertical Pitch Location') +
  ylim(0, 5) +
  facet_wrap(~ game_year)+
  ggtitle('deGrom Fastball Location by Year')

degrom_ff_loc_alt
```

```{r}
syndergaard_cu_loc <- syndergaard_savant_scrape %>%
  filter(pitch_type == 'CU') %>%
  ggplot(aes(plate_x, plate_z)) + 
  geom_hex(bins = 15) + 
  scale_fill_continuous(type = "viridis") +
  theme_bw() +
  geom_path(aes(x, y), data=kZone, lwd=2, col="red") + 
  xlab('Horizontal Pitch Location') +
  ylab('Vertical Pitch Location') + 
  ylim(0, 4.5) +
  facet_wrap(~ game_year) +
  ggtitle('Syndergaard Curveball Location by Year')

syndergaard_cu_loc
```

```{r}
# Syndergaard release point by year
syndergaard_release_point <- syndergaard_savant_scrape %>%
  ggplot(aes(x = release_pos_x, y = release_pos_z, color = pitch_name)) +
  geom_point() +
  xlab('Horizontal Pitch Location') +
  ylab('Vertical Pitch Location') +
  xlim(-2, 2) +
  ylim(0, 7) +
  ggtitle('Syndergaard Release Point by Pitch') +
  facet_wrap(~ game_year)

syndergaard_release_point
```

```{r}
# Bauer release point by year
bauer_release_point <- bauer_savant_scrape %>%
  ggplot(aes(x = release_pos_x, y = release_pos_z, color = pitch_name)) +
  geom_point() +
  xlab('Horizontal Pitch Location') +
  ylab('Vertical Pitch Location') +
  xlim(-2, 2) +
  ylim(0, 7) +
  ggtitle('Bauer Release Point by Pitch') +
  facet_wrap(~ game_year)

bauer_release_point
```

```{r}
# deGrom release point by year
degrom_release_point <- degrom_savant_scrape %>%
  ggplot(aes(x = release_pos_x, y = release_pos_z, color = pitch_name)) +
  geom_point() +
  xlab('Horizontal Pitch Location') +
  ylab('Vertical Pitch Location') +
  xlim(-2, 2) +
  ylim(0, 7) +
  ggtitle('deGrom Release Point by Pitch') +
  facet_wrap(~ game_year)

degrom_release_point
```

```{r}
# Syndergaard xwOBA by year
# Syndergaard
syndergaard_xwoba <- savant_data %>%
  filter(player_name == 'Noah Syndergaard') %>%
  ggplot(aes(x = game_year, y = xwoba, color = pitch_name, group = pitch_name)) +
  geom_line() +
  geom_point(shape = 20, size = 5) +
  xlab('Year') +
  ylab('xwOBA') +
  ggtitle('Syndergaard xwOBA by Pitch & Year')

syndergaard_xwoba
```

```{r}
# Syndergaard wOBA by year
# Syndergaard
syndergaard_woba <- savant_data %>%
  filter(player_name == 'Noah Syndergaard') %>%
  ggplot(aes(x = game_year, y = woba, color = pitch_name, group = pitch_name)) +
  geom_line() +
  geom_point(shape = 20, size = 5) +
  xlab('Year') +
  ylab('wOBA') +
  ggtitle('Syndergaard wOBA by Pitch & Year')

syndergaard_woba
```

